---
title: "Penguins: Day 1"
author: ""
date: "2024-06-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidymodels)
library(dplyr)
library(ggplot2)
```

## Penguin Exploration

Data were collected and made available by Dr. Kristen Gorman and the Palmer Station, Antarctica LTER, a member of the Long Term Ecological Research Network. We are going to explore and visualize these data on day 1!

Hints: In RStudio, you will write code in chunks between the ``` marks. To run code, you will click the green arrow in the far right of the code chunk. Lines 29 - 31 are a code chunk. Variable names may not include blank spaces.

1. Take a glimpse of the data set `penguins`. 

```{r, eval=FALSE}
glimpse() # insert data set name inside the ().
```

How many rows does this data set have? How many columns? What is the variable name in the output that will give you the penguin's flipper length in millimeters?

**Add responses** 

2. Let's calculate a summary statistic! In the code chunk below, fill in the ... to calculate the mean body mass (g) for the Chinstrap species of penguin. 

Hint: Look at your question 1 output to see your variable names!

```{r, eval=FALSE}
penguins |>
  filter(species == "...") |> # insert the name of the penguin species inside the " "
  summarize(mean_bm = mean(...)) # insert the variable name
```


3. Now, create a scatterplot of flipper length on the x-axis, body mass on the y-axis, using species to color the results, by filling in the ... below with correctly formatted variable names. 

Hint: Look at your question 1 output to see your variable names!

```{r, eval=FALSE}
penguins |>
  ggplot(
    aes(x = flipper_length_mm , y = body_mass_g , color = species) # insert your variable names here
  ) + 
  geom_point() +
  labs(x = "flipper length (mm)", 
       y = "body mass (g)") 
```

What are some characteristics you notice?

**Add response**

## Looking forward 

Let's compare body mass and flipper length across a subset of penguins! What can we take away from this graph? The following code chunk will produce charts to help you answer this question.  

```{r, eval=FALSE}
penguins |>
  filter(island %in% c("Torgersen", "Biscoe"),
         !is.na(sex)) |>
  ggplot(
    aes(x = body_mass_g, y = flipper_length_mm, color = species)
  ) + 
  geom_point(alpha = 0.4) + 
  geom_smooth(method = "loess", se = F) +
  facet_wrap(~sex) +  
  labs(
    title = "Penguins body mass vs flipper length",
    subtitle = "on Torgersen and Biscoe island",
    y = "flipper length (mm)",
    x = "body mass (g)",
    color = "species",
    caption = "data and more found here: https://allisonhorst.github.io/palmerpenguins/"
  ) + 
  theme(axis.text.x=element_text(angle=45, hjust=1)) + 
  theme_bw()
```


We will learn and practice with each of the following functions and arguments above as the semester goes on!

Now click "Knit" above to produce a nice HTML document of your results.  

## Example Lessons 

If you would like to see more instances of the container used in class, please see the following class websites and lab-0 instructions used to help students work with containers! 

This is Prof. Mine Çetinkaya-Rundel's Intro to Data Science course: 

https://sta199-s24.github.io/

This is day 1 lab from that course: 

https://sta199-s24.github.io/labs/lab-0.html

This is Dr. Elijah Meyer's Intro to Data Science course: 

https://sta199-f23-1.github.io/


